### 2018-01-11 submission/manuscript.Rmd
update discussion to match better with new direction

Within stool, common resident microbes were most consistently present in the top 10 genera or OTUs across study [Figure 6]. Changes in *Bacteroides*, *Ruminococcaceae*, *Ruminococcus*, and *Roseburia* were consistently found to be disciminative across the different studies for both adenoma and carcinoma [Figure 6]. This data would suggest that whether the non-resident bacterium is *Fusobacteria* or *Peptostreptococcus* is not as important as how these bacteria interact with the changing resident community. Based on these observations, it is possible to hypothesize that tsmall changes in community structure lead to new niches in which any one of the CRC-associated genera can gain a foothold, exacerbating the initial changes in community and facilitating the transition from adenoma to carcinoma.

The tissue studies did not provide a clearer understanding of how the microbiota may be associated with lesion. Generally, the OTU models of unmatched and matched [Figure S7E & F] tissue samples were concordant with stool showing that resident microbes were the most prevalent in the top 10 across study. Unlike in stool, *Fusobacterium* was the only mouth-associated bacteria consistently present in the top 10 of the CRC models [Figure S7B-C & E-F]. The majority of the tissue results seem to be study specific with many top 10 taxa being present only in a single study. Additionally, the presence of genera associated with contamination, within the top 10 most important variables for the genera and OTU models, is worrying because it is commonly a marker of contamination. The low bacterial biomass of tissue samples coupled with potential contaminationcould ultimately explain why the these results seem to be more sporadic than the stool results.

>>>>

Within stool, common GI microbes were most consistently present in the top 10 genera or OTUs across study [Figure 6]. Changes in *Bacteroides*, *Ruminococcaceae*, *Ruminococcus*, and *Roseburia* were consistently found to be disciminative across the different studies for both adenoma and carcinoma [Figure 6]. This data would suggest that whether the non-resident bacterium is *Fusobacteria* or *Peptostreptococcus* is not as important as how these bacteria interact with the changing resident community. Based on these observations, it is possible to hypothesize that small changes in community structure lead to new niches in which any one of the mouth-associated genera can gain a foothold, exacerbating the initial changes in community and facilitating the transition from adenoma to carcinoma.

The tissue studies did not provide a clearer understanding of how the microbiota may be associated with lesion. Generally, the full OTU-based models of unmatched and matched [Figure S7E & F] tissue samples were concordant with stool showing that resident microbes were the most prevalent in the top 10 across study. Unlike in stool, *Fusobacterium* was the only mouth-associated bacteria consistently present in the top 10 of the full CRC models [Figure S7B-C & E-F]. The majority of the tissue results seem to be study specific with many top 10 taxa being present only in a single study. Additionally, the presence of genera associated with contamination, within the top 10 most important variables for the genera and OTU models, is worrying because it is commonly a marker of contamination. The low bacterial biomass of tissue samples coupled with potential contaminationcould ultimately explain why the these results seem to be more sporadic than the stool results.

### 2018-01-11 submission/manuscript.tex
update discussion to match better with new direction

Within stool, common resident microbes were most consistently present in
the top 10 genera or OTUs across study {[}Figure 6{]}. Changes in
\emph{Bacteroides}, \emph{Ruminococcaceae}, \emph{Ruminococcus}, and
\emph{Roseburia} were consistently found to be disciminative across the
different studies for both adenoma and carcinoma {[}Figure 6{]}. This
data would suggest that whether the non-resident bacterium is
\emph{Fusobacteria} or \emph{Peptostreptococcus} is not as important as
how these bacteria interact with the changing resident community. Based
on these observations, it is possible to hypothesize that tsmall changes
in community structure lead to new niches in which any one of the
CRC-associated genera can gain a foothold, exacerbating the initial
changes in community and facilitating the transition from adenoma to
carcinoma.

The tissue studies did not provide a clearer understanding of how the
microbiota may be associated with lesion. Generally, the OTU models of
unmatched and matched {[}Figure S7E \& F{]} tissue samples were
concordant with stool showing that resident microbes were the most
prevalent in the top 10 across study. Unlike in stool,
\emph{Fusobacterium} was the only mouth-associated bacteria consistently
present in the top 10 of the CRC models {[}Figure S7B-C \& E-F{]}. The
majority of the tissue results seem to be study specific with many top
10 taxa being present only in a single study. Additionally, the presence
of genera associated with contamination, within the top 10 most
important variables for the genera and OTU models, is worrying because
it is commonly a marker of contamination. The low bacterial biomass of
tissue samples coupled with potential contaminationcould ultimately
explain why the these results seem to be more sporadic than the stool
results.

>>>>

Within stool, common GI microbes were most consistently present in the
top 10 genera or OTUs across study {[}Figure 6{]}. Changes in
\emph{Bacteroides}, \emph{Ruminococcaceae}, \emph{Ruminococcus}, and
\emph{Roseburia} were consistently found to be disciminative across the
different studies for both adenoma and carcinoma {[}Figure 6{]}. This
data would suggest that whether the non-resident bacterium is
\emph{Fusobacteria} or \emph{Peptostreptococcus} is not as important as
how these bacteria interact with the changing resident community. Based
on these observations, it is possible to hypothesize that small changes
in community structure lead to new niches in which any one of the
mouth-associated genera can gain a foothold, exacerbating the initial
changes in community and facilitating the transition from adenoma to
carcinoma.

The tissue studies did not provide a clearer understanding of how the
microbiota may be associated with lesion. Generally, the full OTU-based
models of unmatched and matched {[}Figure S7E \& F{]} tissue samples
were concordant with stool showing that resident microbes were the most
prevalent in the top 10 across study. Unlike in stool,
\emph{Fusobacterium} was the only mouth-associated bacteria consistently
present in the top 10 of the full CRC models {[}Figure S7B-C \& E-F{]}.
The majority of the tissue results seem to be study specific with many
top 10 taxa being present only in a single study. Additionally, the
presence of genera associated with contamination, within the top 10 most
important variables for the genera and OTU models, is worrying because
it is commonly a marker of contamination. The low bacterial biomass of
tissue samples coupled with potential contaminationcould ultimately
explain why the these results seem to be more sporadic than the stool
results.
